---
title: "ANOVA conventions may not work as you wish"
author: "[Haiyang Jin](https://haiyangjin.github.io/) and Yuzhu Ji"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    code_folding: hide
    number_sections: true
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: paged
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

# Preparations

```{r setup, include=FALSE}
## load libraries
library(knitr)
library(tidyverse)
library(afex)
library(emmeans)
library(vcd)
library(ggpubr)
library(RColorBrewer)

theme_set(papaja::theme_apa())

# set global chunk options, put figures into folder
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  include = TRUE, tidy = FALSE, archive = FALSE,
  size = "big", fig.width=8, fig.asp=0.7, 
  width = 1800,
  emmeans=list(msg.interaction=FALSE))

source(here::here("R", "funcs.R"))
source(here::here("R", "func_kappa.R"))
set.seed(2022)
```

```{r}
two_colors <- c("#D55E00", "#56B4E9")
custom_clr <- c("#7F7F7F", "#E64B35", "#56B4E9", "#DF8F44", "#008B45", "#925E9F", "#415384", "#000000")
```


## Read data

```{r}
# article ID and journal
df_publ <- readxl::read_excel(here::here("data", "CodingSheet_anova.xlsx"),
                              sheet="PaperMeta",
                              range="A1:G750") %>%
  transmute(`Paper ID`, journal = `Publication Title`) %>% 
  mutate(journal = factor(journal),
         journal = fct_recode(journal,
                              JEPG="Journal of Experimental Psychology: General",
                              `Psych Sci`="Psychological Science",
                              `J Abn Psych`="Journal of Abnormal Psychology",
                              JCCP="Journal of Consulting and Clinical Psychology",
                              JESP="Journal of Experimental Social Psychology",
                              JPSP="Journal of Personality and Social Psychology"),
         journal = fct_relevel(journal, "JEPG", "Psych Sci", "J Abn Psych", "JCCP", "JESP", "JPSP")) 

# "Title_Prescreening" Sheet
df_prescreen <- readxl::read_excel(here::here("data", "CodingSheet_anova.xlsx"),
                                   sheet="Title_Prescreening",
                                   range="A1:C750") %>%
  left_join(df_publ, by="Paper ID") %>%
  select(-Title)

# "ResearchPaper" Sheet
df_mainscreen <- readxl::read_excel(here::here("data", "CodingSheet_anova.xlsx"),
                               sheet="ResearchPaper",
                               range="A1:L750") %>%
  left_join(df_publ, by="Paper ID") %>%
  select(-Title)
```

```{r}
# read the coding sheet for the first round by YJ
# df_conv_YJ1 <- readxl::read_excel(here::here("data", "CodingSheet_pilot.xlsx"), 
#                                  sheet="AnalysisInfo_YJ",
#                                  range="A1:AO39") %>% 
#   select(-Title) 
# 
# # read the coding sheet for the first round by HJ
# df_conv_HJ1 <- readxl::read_excel(here::here("data", "CodingSheet_pilot.xlsx"), 
#                                  sheet="AnalysisInfo_HJ",
#                                  range="A1:AO39") %>% 
#   select(-Title) 
```


```{r}
# # read the coding sheet for the first round (after resolving inconsistency)
# df_conv_1 <- readxl::read_excel(here::here("data", "CodingSheet_pilot.xlsx"), 
#                                  sheet="AnalysisInfo_1st",
#                                  range="A1:AO39") %>% 
#   select(-Title) 
# # read the coding sheet for the second round by YJ
# df_conv_YJ2 <- readxl::read_excel(here::here("data", "CodingSheet_pilot.xlsx"), 
#                                  sheet="AnalysisInfo_2nd_YJ",
#                                  range="A1:AO39") %>% 
#   select(-Title) 
# # read the coding sheet for the second round by YJ
# df_conv_HJ2 <- readxl::read_excel(here::here("data", "CodingSheet_pilot.xlsx"), 
#                                  sheet="AnalysisInfo_2nd_YJ",
#                                  range="A1:AO39") %>% 
#   select(-Title) 
# 
# df_conv <- bind_rows(df_conv_1, df_conv_YJ2, df_conv_HJ2) %>% 
#   left_join(df_publ, by="Paper ID")

# df_conv <- readxl::read_excel(here::here("data", "CodingSheet_pilot.xlsx"), 
#                                  sheet="AnalysisInfo_HJ",
#                                  range="A1:AO39") %>% 
#   left_join(df_publ, by="Paper ID") %>% 
#   select(-Title) 
```


# Intra-rater consistency
Estimate the intra-rater consistency for the information on whether the study employed ANOVA, experimental designs and the ANOVA prevalence inspected by Y.J. and H.J. separately.

## Research article and ANOVA

```{r}
# # whether it is a research article (unweighted)
df_res_K <- df_mainscreen %>%
  select(`Research article_YJ`, `Research article_HJ`) %>%
  group_by(`Research article_YJ`, `Research article_HJ`) %>%
  summarize(count = n(), .groups = "drop") %>%
  pivot_wider(names_from=`Research article_HJ`, values_from = count, values_fill = 0)

Ka_res <- df_res_K %>%
  select(-`Research article_YJ`) %>%
  as.matrix() %>%
  Kappa()

print(Ka_res, CI=TRUE) # Kappa for research articles
```

```{r}
# whether ANOVA was used (unweighted)
df_aov_K <- df_mainscreen %>% 
  filter(`Research article`==1) %>% # only among research articles
  select(`Use ANOVA_YJ`, `Use ANOVA_HJ`) %>% 
  group_by(`Use ANOVA_YJ`, `Use ANOVA_HJ`) %>% 
  summarize(count=n(), .groups="drop") %>% 
  pivot_wider(names_from=`Use ANOVA_HJ`, values_from = count, values_fill = 0) 

Ka_aov <- df_aov_K %>% 
  select(-`Use ANOVA_YJ`) %>% 
  as.matrix() %>% 
  Kappa()

print(Ka_aov, CI=TRUE) # Kappa for using ANOVA
```


## The first 120 articles

```{r}
# experimental designs
# design_list <- c("N_factor", "max_No_levels")
# 
# kappa_0 <- lapply(design_list, thekappa, df1=df_conv_YJ1, df2=df_conv_HJ1, isweighted=TRUE) %>% 
#   bind_rows()
```

```{r}
# Convention A
# con_a_list <- c("a1_report_all", "a3_correction")
# 
# kappa_a <- lapply(con_a_list, thekappa, 
#                   df1=filter(df_conv_YJ1, a2_multiway==1), 
#                   df2=filter(df_conv_HJ1, a2_multiway==1)) %>% 
#   bind_rows()
```

```{r}
# Convention B
# con_b_wlist <- c("b1_N_factor3+", "b2_N_significant_b1")
# con_b_ulist <- c("b7_posthoc", "b8_correction")
# 
# kappa_bw <- lapply(con_b_wlist, thekappa, isweighted=TRUE,
#                    df1=filter(df_conv_YJ1, max_No_levels>2), 
#                    df2=filter(df_conv_HJ1, max_No_levels>2)) %>% 
#   bind_rows()
# 
# kappa_bu <- lapply(con_b_ulist, thekappa, 
#                    df1=filter(df_conv_YJ1, max_No_levels>2), 
#                    df2=filter(df_conv_HJ1, max_No_levels>2)) %>% 
#   bind_rows()
# kappa_b <- bind_rows(kappa_bw, kappa_bu)
```

```{r}
# Convention C
# con_c_wlist <- c("c1_N_2way_inter", "c2_N_sig_in_c1")
# con_c_ulist <- c("c7_simple_effect", "c8_correction")
# 
# kappa_cw <- lapply(con_c_wlist, thekappa, isweighted=TRUE,
#                    df1=filter(df_conv_YJ1, N_factor>1), 
#                    df2=filter(df_conv_HJ1, N_factor>1)) %>% 
#   bind_rows()
# 
# kappa_cu <- lapply(con_c_ulist, thekappa, 
#                    df1=filter(df_conv_YJ1, N_factor>1), 
#                    df2=filter(df_conv_HJ1, N_factor>1)) %>% 
#   bind_rows()
# kappa_c <- bind_rows(kappa_cw, kappa_cu)
```

```{r}
# Convention D & E
# kappa_d1e1 <- thekappa(df1=filter(df_conv_YJ1, N_factor>2), 
#                     df2=filter(df_conv_HJ1, N_factor>2),
#                     colname="d1_e1_highest_sig")
# # kappa_d <- thekappa(df1=filter(df_conv_YJ1, N_factor>2, d1_e1_highest_sig==1), 
# #                     df2=filter(df_conv_HJ1, N_factor>2, d1_e1_highest_sig==1),
# #                     colname="d1_separate")
# kappa_e <- thekappa(df1=filter(df_conv_YJ1, N_factor>2, d1_e1_highest_sig==0), 
#                     df2=filter(df_conv_HJ1, N_factor>2, d1_e1_highest_sig==0),
#                     colname="e1_combine")
# 
# # kappa_de <- bind_rows(kappa_d1e1, kappa_d, kappa_e)
# kappa_de <- bind_rows(kappa_d1e1, kappa_e)
```

```{r}
# (kappa_df <- bind_rows(kappa_0, kappa_a, kappa_b, kappa_c, kappa_de))
```

```{r}
# the mean consistency
# tibble(mean = mean(kappa_df$value),
#        min = min(kappa_df$value),
#        max = max(kappa_df$value)) 
```

# ANOVA prevalence

## Screening information

```{r}
# number of articles included at each stage
df_included_all <- df_mainscreen %>%  
  summarize(journal = "Overall",
            In1_all = n(),
            In2_title = sum(1-Title_Prescreening),
            In3_research = sum(`Research article`),
            In4_ANOVA = sum(`Use ANOVA`),
            In5_include = sum(Included))

# number of articles at different stages
df_included <- df_mainscreen %>% 
  group_by(journal) %>% 
  summarize(In1_all = n(),
            In2_title = sum(1-Title_Prescreening),
            In3_research = sum(`Research article`),
            In4_ANOVA = sum(`Use ANOVA`),
            In5_include = sum(Included)) %>% 
  add_row(df_included_all, .before=1)
df_included
```

```{r}
# Percentage (relative to the total number of articles)
df_included %>% 
  transmute(journal, In1_all,
            In2_title = paste0(round((In2_title/In1_all)*100, 2), '%'),
            In3_research = paste0(round((In3_research/In1_all)*100, 2), '%'),
            In4_ANOVA = paste0(round((In4_ANOVA/In1_all)*100, 2), '%'),
            In5_include = paste0(round((In5_include/In1_all)*100, 2), '%'))
```


```{r}
# number of articles excluded at different stages
df_excluded <- df_mainscreen %>% 
  group_by(journal) %>% 
  summarize(Ex1_all = n(),
            Ex2_title = sum(Title_Prescreening),
            Ex3_research = sum(1-`Research article`),
            Ex4_ANOVA = sum(1-`Use ANOVA`),
            Ex5_exclude = sum(1-Included)) %>% 
  add_row( # add summary row
    summarize(df_mainscreen, 
              journal = "Overall",
              Ex1_all = n(),
              Ex2_title = sum(Title_Prescreening),
              Ex3_research = sum(1-`Research article`),
              Ex4_ANOVA = sum(1-`Use ANOVA`),
              Ex5_exclude = sum(1-Included)),
    .before=1)
df_excluded
```

## ANOVA prevalence
```{r}
# The percentage of articles employing ANOVA among all research articles:  
df_included %>% 
  mutate(prevalence_anova = round(In4_ANOVA/In3_research*100, 2))
```

## ANOVA complexity

```{r}
# df_complex <- df_conv %>% 
#   filter(included==1) %>% 
#   transmute(`Paper ID`, N_factor, max_No_levels, Design,
#             Design_int = str_split(Design, "_"),
#             Avg_No_levels = sapply(Design_int, function(x){mean(as.numeric(x))}),
#             Complexity = sapply(Design_int, function(x){prod(as.numeric(x))}),
#             journal)
```

Across ANOVAs:
```{r}
# df_complex %>% 
#   group_by(journal) %>% 
#   summarize(mean_N_factor = mean(N_factor),
#             mean_max_N_levels = mean(max_No_levels),
#             mean_complexity = mean(Complexity)) %>% 
#   add_row( # add summary row
#     summarize(df_complex, 
#               journal = "All",
#               mean_N_factor = mean(N_factor),
#               mean_max_N_levels = mean(max_No_levels),
#               mean_complexity = mean(Complexity))) 
```

```{r}
# plot_design <- df_complex %>% 
#   mutate(design_ = str_replace_all(Design, '_', '*')) %>% 
#   ggplot(aes(x = reorder(design_, Complexity))) +
#   geom_histogram(stat="count") +
#   labs(x = "Experimental designs")
# plot_design
```

alluvial figure (?) [left side is the experimental designs and the right side is the numbers of conditions.]
```{r}
# df_complex %>% 
#   select(Design, Complexity) %>% 
#   group_by(Design, Complexity) %>% 
#   summarize(count = n(), .groups="drop") 
```

## ANOVA convention prevalence

### Convention A: inspect all effects
```{r}
# df_conv_a <- df_conv %>% 
#   filter(a2_multiway==1) %>% 
#   select(`Paper ID`, starts_with("a"), journal)
# 
# # percentage of ANOVAs reporting/inspecting all effects among multi-way ANOVA
# df_conv_a_all <- summarize(df_conv_a, 
#                            journal = "All",
#                            conA = mean(a1_report_all))
# 
# df_conv_a %>% 
#   group_by(journal) %>% 
#   summarize(conA = mean(a1_report_all)) %>% 
#   add_row(df_conv_a_all) # add summary row
```

```{r}
# df_conv_a_adjust <- df_conv_a %>% 
#   group_by(a3_correction) %>% 
#   summarize(journal = "All", count = n())
# 
# df_conv_a %>% 
#   group_by(journal, a3_correction) %>% 
#   summarize(count = n(), .groups="drop") %>% 
#   add_row(df_conv_a_adjust) 
```


### Convention B: post-hoc
```{r}
# df_conv_b <- df_conv %>% 
#   filter(max_No_levels>2) %>% 
#   select(`Paper ID`, max_No_levels, starts_with("b"), journal)
# 
# # percentage of (not) performing post-hoc analysis when the main effect is (not) significant
# df_conv_b_artile <- df_conv_b %>% 
#   mutate(isConB = b5_main_effect_sig == b7_posthoc) %>% 
#   group_by(`Paper ID`, journal) %>% 
#   summarize(meanB_art = mean(isConB), .groups="drop")
# 
# # show the result
# df_conv_b_all <- df_conv_b_artile %>% 
#   summarize(journal = "All", conB = mean(meanB_art))
# 
# df_conv_b_artile %>% 
#   group_by(journal) %>% 
#   summarize(conB = mean(meanB_art)) %>% 
#   add_row(df_conv_b_all) # add summary row
```

Multiple comparison corrections (per ANOVA)
```{r}
# df_conv_b_adjust <- df_conv_b %>% 
#   filter(b5_main_effect_sig == 1) %>% 
#   group_by(b8_correction) %>% 
#   summarize(journal = 'All', count = n())
#   
# df_conv_b %>% 
#   filter(b5_main_effect_sig == 1) %>% 
#   group_by(journal, b8_correction) %>% 
#   summarize(count = n(), .groups="drop") %>% 
#   add_row(df_conv_b_adjust) 
```


### Convention C: simple effect analysis
```{r}
# df_conv_c <- df_conv %>% 
#   filter(N_factor>1) %>% 
#   select(`Paper ID`, N_factor, starts_with("c") & !starts_with("Con"), journal)
# 
# # percentage of (not) performing simple analysis when the interaction is (not) significant
# df_conv_c_artile <- df_conv_c %>% 
#   mutate(isConC = c5_inter_sig == c7_simple_effect) %>% 
#   group_by(`Paper ID`, journal) %>% 
#   summarize(meanC_art = mean(isConC), .groups="drop")
# 
# # show the result
# df_conv_c_all <- df_conv_c_artile %>% 
#   summarize(journal = "All", conC = mean(meanC_art))
# 
# df_conv_c_artile %>% 
#   group_by(journal) %>% 
#   summarize(conC = mean(meanC_art)) %>% 
#   add_row(df_conv_c_all) # add summary row
```

Multiple comparison corrections (per ANOVA)
```{r}
# df_conv_c_adjust <- df_conv_c %>% 
#   filter(c5_inter_sig == 1) %>% 
#   group_by(c8_correction) %>% 
#   summarize(journal = 'All', count = n())
#   
# df_conv_c %>% 
#   filter(c5_inter_sig == 1) %>% 
#   group_by(journal, c8_correction) %>% 
#   summarize(count = n(), .groups="drop") %>% 
#   add_row(df_conv_c_adjust)
```

### Convention D: separate 

```{r}
# df_conv_de <- df_conv %>% 
#   filter(N_factor>2) %>% 
#   select(`Paper ID`, N_factor, d1_e1_highest_sig, d1_separate, e1_combine, journal)
# 
# # 
# df_conv_d_artile <- df_conv_de %>% 
#   filter(d1_e1_highest_sig == 1) %>% 
#   group_by(`Paper ID`, journal) %>% 
#   summarize(meanD_art = mean(d1_separate), .groups="drop")
# 
# # show the result
# df_conv_d_all <- df_conv_d_artile %>% 
#   summarize(journal = "All", conD = mean(meanD_art))
# 
# df_conv_d_artile %>% 
#   group_by(journal) %>% 
#   summarize(conD = mean(meanD_art)) %>% 
#   add_row(df_conv_d_all) # add summary row
```

### Convention E: combine 

```{r}
# # 
# df_conv_e_artile <- df_conv_de %>% 
#   filter(d1_e1_highest_sig == 0) %>% 
#   group_by(`Paper ID`, journal) %>% 
#   summarize(meanE_art = mean(e1_combine), .groups="drop")
# 
# # show the result
# df_conv_e_all <- df_conv_e_artile %>% 
#   summarize(journal = "All", conE = mean(meanE_art))
# 
# df_conv_e_artile %>% 
#   group_by(journal) %>% 
#   summarize(conE = mean(meanE_art)) %>% 
#   add_row(df_conv_e_all) # add summary row
```


# Simulating Type I error in ANOVA

```{r}
Nsim <- 10000 # number of simulation
alphas <- c(0.001, 0.005, 0.01, 0.05, (1:4)/10)  # 0.001 ~ 0.4
```

## ANOVA omnibus F-test

```{r, message=F}
p_omni <- sim_omnibus(N_subj = 30, iter = Nsim, n_core=16, 
                      file_cache = here::here("simulation", "p_omni.rds"),
                      N_IV = 2:5)
```

### FWER and CER

```{r}
# Type I error rates in different situations
df_TypeI_omni <- sig_omnibus(p_omni, 0.05) %>% 
  select(-c(alpha, effnames, p.value, sig, omniF)) %>% 
  distinct() %>% 
  group_by(N_IV, adjust) %>% 
  summarize(TypeI_omniF = mean(sig_omniF), 
            TypeI_any = mean(sig_any),
            `TypeI_omniF&any` = mean(`sig_omni&any`),
            .groups = "drop") 

df_TypeI_omni
```

```{r fig.width=4, fig.asp=1.2}
# Conclusion-based Type I error rate (CBER)
df_omni_cber <- df_TypeI_omni %>% 
  pivot_longer(starts_with("TypeI_"), names_to = "effnames", values_to = "TypeI") %>% 
  mutate(adjust = if_else(effnames == "TypeI_omniF", "uncorrected", adjust),
         adjust = fct_relevel(adjust, "uncorrected"),
         effnames = factor(effnames, 
                           levels = c("TypeI_omniF", "TypeI_any", "TypeI_omniF&any")),
         effnames = fct_recode(effnames, `omnibus\nF-test` = "TypeI_omniF", 
                              `Inspect all\n(FWER)`= "TypeI_any", 
                              `omnibus & all\n(CER)`="TypeI_omniF&any")) %>% 
  distinct() 

fig_simu_a <- df_omni_cber %>% 
  filter(N_IV == 2) %>% # only show results for ANOVAs with two factors
  ggplot(aes(x=as.numeric(adjust), y = TypeI*100, fill=adjust)) + 
  geom_col(width=1) +   
  facet_grid(cols=vars(effnames), scales="free_x", space="free_x", switch="x")+
  ggh4x::facetted_pos_scales(
    list(
      scale_x_continuous(breaks = 1, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 1.5, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 1.5, expand = c(0, 0.65))
    )) + 
  scale_fill_manual(values = custom_clr[c(1,5)],
                    breaks = levels(df_omni_cber$adjust)) + #,  
  geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
  scale_y_continuous(limits = c(0, 24), expand = c(0, 0)) +
  labs(y = "Type I Error Rate (%)", fill = "") +
  theme(legend.position = c(0.75, 0.6),
        axis.ticks.length = unit(.15, "cm"),
        axis.text.x = element_blank(),
        panel.spacing.x = unit(0, "pt"),
        strip.placement = "outside",
        axis.title.x = element_blank())
  
# ggsave(file.path("figures","fig_simu_TypeI_omni.png"), fig_simu_a, width = 4, height = 5.6)
fig_simu_a
```

```{r fig.width=14, fig.asp=0.4}
# show results for ANOVAs with different factor numbers
for (i in unique(df_omni_cber$N_IV)) {
  
  nam <- paste("fig_simu_a_tmp", i, sep = "")
  
  assign(
    nam, df_omni_cber %>% 
      filter(N_IV == i) %>% 
      ggplot(aes(x=as.numeric(adjust), y = TypeI*100, fill=adjust)) + 
      geom_col(width=1) +   
      facet_grid(cols=vars(effnames), scales="free_x", space="free_x", switch="x")+
      ggh4x::facetted_pos_scales(
        list(
          scale_x_continuous(breaks = 1, expand = c(0, 0.65)),
          scale_x_continuous(breaks = 1.5, expand = c(0, 0.65)),
          scale_x_continuous(breaks = 1.5, expand = c(0, 0.65))
        )) + 
      scale_fill_manual(values = custom_clr[c(1,5)],
                        breaks = levels(df_omni_cber$adjust)) + #,  
      geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
      scale_y_continuous(limits = c(0, 84), expand = c(0, 0)) +
      labs(x = paste0("ANOVAs with ", i, " factors"), y = "Type I Error Rate (%)", fill = "") +
      theme(legend.position = c(0.76, 0.6),
            axis.ticks.length = unit(.15, "cm"),
            axis.text.x = element_blank(),
            panel.spacing.x = unit(0, "pt"),
            strip.placement = "outside",
            axis.title.x = element_text(margin = margin(t = 0)))
  )
}

fig_simu_a_supp <- ggarrange(fig_simu_a_tmp2, fig_simu_a_tmp3, 
                             fig_simu_a_tmp4, fig_simu_a_tmp5, 
                             nrow = 1)
  
# ggsave(file.path("figures","fig_simu_TypeI_omni_supp.png"), fig_simu_a_supp, width = 14, height = 5.6)
fig_simu_a_supp
```


### Single tests

```{r}
# with omnibus F-test and significant effect (uncorrected)
df_sig_single1 <- sig_omnibus(p_omni, 0.05)  %>% 
  filter(adjust == "uncorrected") %>% 
  transmute(Method = "omnibus F-test",
            N_IV, effnames, 
            sig_cber = `sig_omni&any`,
            sig_cber_single = sig_cber & sig)

df_TypeI_single <- sig_omnibus(p_omni, 0.05)  %>% 
  filter(adjust == "Bonferroni") %>% 
  transmute(Method = "Bonferroni",
            N_IV, effnames,
            sig_cber = sig_any,
            sig_cber_single = sig_cber & sig) %>% 
  # combine the two df and make it to long format
  bind_rows(df_sig_single1) %>% 
  group_by(N_IV, effnames, Method) %>% 
  summarize(TypeI_cber = mean(sig_cber), 
            TypeI_cber_single = mean(sig_cber_single),
            .groups = "drop") %>% 
  pivot_wider(c(N_IV, Method, TypeI_cber), 
              names_from = effnames, values_from = TypeI_cber_single) %>% 
  pivot_longer(!c("N_IV", "Method"), 
               names_to = "effnames", values_to = "TypeI") %>% 
  filter(!is.na(TypeI)) %>% 
  # update the level names and order
  mutate(Method = factor(Method, levels=c("omnibus F-test", "Bonferroni")))

# Type I error rate for single tests (when the CBER is controlled)
df_TypeI_single %>%
  pivot_wider(names_from = effnames, values_from = TypeI) %>% 
  select(Method, N_IV, TypeI_cber, everything())
```

```{r fig.width=4, fig.asp=1.2}
fig_simu_single_a <- df_TypeI_single %>% 
  filter(N_IV == 2) %>% 
  select(-c(N_IV)) %>% 
  mutate(effnames = factor(effnames),
         effnames = fct_relevel(effnames, "TypeI_cber", "A", "B"),
         effnames = fct_recode(effnames, CBER="TypeI_cber", `main effect A`="A",
                               `main effect B`="B", interaction="A:B"),
         Method = paste0(Method, "\n"),
         Method = factor(Method, levels=c("omnibus F-test\n", "Bonferroni\n"))) %>% 
  ggplot(aes(x = Method, y = TypeI*100, fill = effnames, group = effnames)) +
  geom_col(position = "dodge") +
  # facet_grid(~ omni, switch = "x") + 
  scale_fill_manual(values = custom_clr[1:4]) +
  geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
  scale_y_continuous(limits = c(0, 7), expand = c(0, 0), breaks = 1:6) +
  labs(y = "Type I Error Rate (%)", fill = "") +
  guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme(legend.position = c(0.55, 0.87),
        legend.direction="horizontal",
        axis.ticks.length = unit(.15, "cm"),
        strip.placement = "outside",
        axis.title.x = element_blank())

# ggsave(file.path("figures","fig_simu_single_omni.png"), fig_simu_single_a, width = 4, height = 5.6)
fig_simu_single_a
```

```{r fig.width=9, fig.asp=0.8}
df_TypeI_single_plot <- df_TypeI_single %>% 
  mutate(effnames = factor(effnames),
         effnames = fct_relevel(effnames, "TypeI_cber"),
         effnames = fct_recode(effnames, CBER="TypeI_cber"),
         Method = as_factor(Method)) 

for (i in unique(df_TypeI_single_plot$N_IV)) {
  
  nam <- paste("fig_simu_single_a_supp_tmp", i, sep = "")
  
  assign(
    nam, df_TypeI_single_plot %>% 
      filter(N_IV==i) %>% 
      ggplot(aes(x = Method, y = TypeI*100, fill = effnames, group = effnames)) +
      geom_col(position = "dodge") +
      # facet_wrap(vars(N_IV), switch = "x") +
      scale_fill_manual(values = c(custom_clr[c(1,3)], brewer.pal(8,"Set1"), 
                                   brewer.pal(7, "Set2"), brewer.pal(8,"Set3"), brewer.pal(7,"Accent"))) + 
      geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
      scale_y_continuous(limits = c(0, 7), expand = c(0, 0), breaks = 1:6) +
      labs(x = paste0("ANOVAs with ", as.character(i), " factors"), y = "Type I Error Rate (%)", fill = "") +
      guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
      theme(legend.position = "none",
            # legend.direction="horizontal",
            axis.ticks.length = unit(.15, "cm"),
            strip.placement = "outside",
            axis.title.x = element_text(margin = margin(t = 5)))
  )
}

fig_simu_single_a_supp <- ggarrange(
  ggarrange(fig_simu_single_a_supp_tmp2, fig_simu_single_a_supp_tmp5,
            ncol=2, widths = c(0.3, 0.7)), 
  ggarrange(fig_simu_single_a_supp_tmp3, fig_simu_single_a_supp_tmp4,
            ncol=2, widths = c(0.4, 0.6)),
  nrow = 2)

# ggsave(file.path("figures","fig_simu_single_omni_supp.png"), fig_simu_single_a_supp, width = 10, height = 7)
fig_simu_single_a_supp
```

## Main effect and post-hoc analysis

```{r message=F}
# sim_main_posthoc is available in R/funcs.R
p_main_posthoc <- sim_main_posthoc(N_subj = 30, iter = Nsim, n_core=16, 
                                   file_cache = here::here("simulation", "p_main_posthoc.rds"),
                                   N_levels = c(3,4,5,6)) 
```

### FWER and CER
```{r}
df_TypeI_posthoc <- sig_main_posthoc(p_main_posthoc, 0.05) %>% 
  group_by(N_level, alpha, adjust) %>% 
  summarize(TypeI_main = mean(sig_main),
            TypeI_postonly = mean(sig_anypost),
            `TypeI_main&posthoc` = mean(`sig_main&posthoc`), 
            .groups = "drop")

df_TypeI_posthoc
```

Claim significant results only when both the main effect and at least one of the post-hoc tests (with different multiple comparison corrections) are significant:
```{r}
df_TypeI_posthoc %>% 
  filter(adjust != "uncorrected") %>% # only when using multiple comparison correction
  select(-c(TypeI_main, TypeI_postonly)) %>% 
  pivot_wider(c(N_level, alpha), 
              names_from=adjust, values_from=`TypeI_main&posthoc`)
```

Claim significant results only when the main effect and at least one of the post-hoc tests (without multiple comparison corrections) are significant:
```{r}
df_TypeI_posthoc %>% 
  filter(adjust == "uncorrected") %>%  # only when NOT using multiple comparison correction
  select(-c(TypeI_main, TypeI_postonly)) %>% 
  pivot_wider(c(N_level, alpha), 
              names_from=adjust, values_from=`TypeI_main&posthoc`)
```

Claim significant results only when at least one of the post-hoc tests (with different multiple comparison corrections) are significant (regardless of the main effect results): 
```{r}
df_TypeI_posthoc %>% 
  filter(adjust != "uncorrected") %>% # only when using multiple comparison correction
  select(-c(TypeI_main, `TypeI_main&posthoc`)) %>% 
  pivot_wider(c(N_level, alpha), 
              names_from=adjust, values_from=TypeI_postonly)
```

```{r fig.width=8, fig.asp=0.7}
df_posthoc_tmp <- df_TypeI_posthoc %>% 
  pivot_longer(starts_with("TypeI_"), names_to = "effects", values_to = "TypeI") %>% 
  mutate(adjust = if_else(effects == "TypeI_main", "uncorrected", 
                          if_else(adjust == "uncorrected", "uncorrected", 
                                  str_to_title(as.character(adjust)))),
         adjust = fct_reorder(adjust, TypeI, mean, .desc=TRUE),
         adjust = fct_relevel(adjust, "uncorrected"),
         effects = factor(effects, 
                          levels = c("TypeI_main", "TypeI_postonly", "TypeI_main&posthoc")),
         effects = fct_recode(effects, `Main effect\n` ="TypeI_main", 
                              `Post-hoc analysis\n(six tests, FWER)`="TypeI_postonly", 
                              `Main effect & Post-hoc analysis\nwith six tests (CER)`="TypeI_main&posthoc")) %>% 
  distinct() 

fig_simu_b <- df_posthoc_tmp %>% 
  filter(N_level == 4) %>% 
  ggplot(aes(x=as.numeric(adjust), y = TypeI*100, fill=adjust)) + 
  geom_col(width=1) +   
  facet_grid(cols=vars(effects), scales="free_x", space="free_x", switch="x")+
  ggh4x::facetted_pos_scales(
    list(
      scale_x_continuous(breaks = 1, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3.5, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3.5, expand = c(0, 0.65))
    )) + 
  scale_fill_manual(values = custom_clr,
                    breaks = levels(df_posthoc_tmp$adjust)) + #,  
  geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
  scale_y_continuous(limits = c(0, 24), expand = c(0, 0)) +
  labs(y = "Type I Error Rate (%)", fill = "") +
  theme(legend.position = c(0.74, 0.63),
        axis.ticks.length = unit(.15, "cm"),
        axis.text.x = element_blank(),
        panel.spacing.x = unit(0, "pt"),
        strip.placement = "outside",
        axis.title.x = element_blank())
  
# ggsave(file.path("figures","fig_simu_TypeI_mainpost.png"), fig_simu_b, width = 8, height = 5.6)
fig_simu_b
```



```{r fig.width=13, fig.asp=.7}

for (i in unique(df_posthoc_tmp$N_level)) {
  
  nam <- paste("fig_simu_b_tmp", i, sep = "")
  
  assign(
    nam, df_posthoc_tmp %>% 
    filter(N_level == i) %>% 
    ggplot(aes(x=as.numeric(adjust), y = TypeI*100, fill=adjust)) + 
    geom_col(width=1) +   
    facet_grid(cols=vars(effects), scales="free_x", space="free_x", switch="x")+
    ggh4x::facetted_pos_scales(
      list(
        scale_x_continuous(breaks = 1, expand = c(0, 0.65)),
        scale_x_continuous(breaks = 3.5, expand = c(0, 0.65)),
        scale_x_continuous(breaks = 3.5, expand = c(0, 0.65))
      )) + 
    scale_fill_manual(values = custom_clr,
                      breaks = levels(df_posthoc_tmp$adjust)) + #,  
    geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
    scale_y_continuous(limits = c(0, 45), expand = c(0, 0)) +
    labs(x = paste0("Main effects with ", i, " levels"), y = "Type I Error Rate (%)", fill = "") +
    theme(legend.position = c(0.75, 0.6),
          axis.ticks.length = unit(.15, "cm"),
          axis.text.x = element_blank(),
          panel.spacing.x = unit(0, "pt"),
          strip.placement = "outside",
          axis.title.x = element_text(margin = margin(t = 0)))
  )
  
}

fig_simu_b_supp <- ggarrange(fig_simu_b_tmp3, fig_simu_b_tmp4, 
                             fig_simu_b_tmp5, fig_simu_b_tmp6, 
                             nrow = 2, ncol = 2)

# ggsave(file.path("figures","fig_simu_TypeI_mainpost_supp.png"), fig_simu_b_supp, width = 13, height = 8)
fig_simu_b_supp
```

### Single tests

```{r}
# with omnibus F-test and significant effect
df_sig_main_single1 <- sig_main_posthoc(p_main_posthoc, 0.05) %>% 
  filter(adjust == "uncorrected") %>% 
  transmute(Method = "Main effect & Post-hoc",
            N_level, contrast,
            sig_cber = `sig_main&posthoc`, # main and post-hoc (uncorrected)
            sig_cber_single = sig_cber & sig)

df_TypeI_main_single <- sig_main_posthoc(p_main_posthoc, 0.05) %>% 
  filter(adjust != "uncorrected") %>% 
  transmute(Method = str_to_title(adjust),
            N_level, contrast,
            sig_cber = `sig_anypost`, # post-hoc only (corrected)
            sig_cber_single = sig_cber & sig) %>% 
  # combine two df and make it to long format
  bind_rows(df_sig_main_single1) %>% 
  group_by(N_level, contrast, Method) %>% 
  summarize(TypeI_cber = mean(sig_cber), # correction and at least one sig
            TypeI_cber_single = mean(sig_cber_single),
            .groups = "drop") %>% 
  pivot_wider(c(N_level, Method, TypeI_cber), 
              names_from = contrast, values_from = TypeI_cber_single) %>% 
  pivot_longer(!c("N_level", "Method"), 
               names_to = "contrast", values_to = "TypeI") %>% 
  filter(!is.na(TypeI))
  
# Type I error rate for single tests (when the FWER is controlled)
df_TypeI_main_single %>%
  pivot_wider(names_from = contrast, values_from = TypeI) %>% 
  select(Method, N_level, TypeI_cber, everything())
```

```{r fig.width=12, fig.asp=0.4}
fig_simu_single_b <- df_TypeI_main_single %>% 
  filter(N_level == 4) %>% 
  select(-c(N_level)) %>% 
  mutate(contrast = factor(contrast),
         contrast = fct_relevel(contrast, "TypeI_cber"),
         contrast = fct_recode(contrast, CBER="TypeI_cber"),
         contrast = str_replace_all(contrast, "IV", "level"),
         Method = if_else(str_starts(Method, "Main"),
                          paste0(Method, "\n(uncorrected)"),
                          paste0(Method, "\n(post-hoc only)")),
         Method = as_factor(Method),
         Method = fct_reorder(Method, TypeI, mean, .desc=TRUE)) %>%
  ggplot(aes(x = Method, y = TypeI*100, fill = contrast, group = contrast)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = custom_clr) +
  geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
  scale_y_continuous(limits = c(0, 7), expand = c(0, 0), breaks = 1:6) +
  labs(y = "Type I Error Rate (%)", fill = "") +
  guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme(legend.position = c(0.5, 0.87),
        legend.direction="horizontal",
        axis.ticks.length = unit(.15, "cm"),
        strip.placement = "outside",
        axis.title.x = element_blank())

# ggsave(file.path("figures","fig_simu_single_mainpost.png"), 
#        fig_simu_single_b, width = 12, height = 4.8)
fig_simu_single_b
```


```{r fig.width=14, fig.asp=1.1}

for (i in unique(df_TypeI_main_single$N_level)) {
  
  nam <- paste("fig_simu_single_b_tmp", i, sep = "")
  
  assign(
    nam, df_TypeI_main_single %>% 
      filter(N_level == i) %>% 
      select(-c(N_level)) %>% 
      mutate(contrast = factor(contrast),
             contrast = fct_relevel(contrast, "TypeI_cber"),
             contrast = fct_recode(contrast, CBER="TypeI_cber"),
             contrast = str_replace_all(contrast, "IV", "level"),
             Method = if_else(str_starts(Method, "Main"), 
                          paste0(Method, "\n(uncorrected)"),
                          paste0(Method, "\n(post-hoc only)")),
             Method = as_factor(Method),
             Method = fct_reorder(Method, TypeI, mean, .desc=TRUE)) %>%
      ggplot(aes(x = Method, y = TypeI*100, fill = contrast, group = contrast)) +
      geom_col(position = "dodge") +
      scale_fill_manual(values = c(custom_clr[c(1,3)], brewer.pal(8,"Set1"),
                                   brewer.pal(6,"Set3"))) +
      geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
      scale_y_continuous(limits = c(0, 7), expand = c(0, 0), breaks = 1:6) +
      labs(x = paste0("Main effects with ", i, " levels"), y = "Type I Error Rate (%)", fill = "") +
      guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
      theme(legend.position = c(0.5, 0.88),
            legend.direction="horizontal",
            axis.ticks.length = unit(.15, "cm"),
            strip.placement = "outside")
  )
}

fig_simu_single_b_supp <- ggarrange(fig_simu_single_b_tmp3, 
                                    fig_simu_single_b_tmp4, 
                                    ncol=2) %>% 
  ggarrange(fig_simu_single_b_tmp5,
            fig_simu_single_b_tmp6,
            ncol=1)

# ggsave(file.path("figures","fig_simu_single_mainpost_supp.png"), 
#        fig_simu_single_b_supp, width = 14, height = 14)

fig_simu_single_b_supp
```


## Interaction and simple effect analysis

```{r, message=F}
p_inter_simple <- sim_inter_simple(N_subj = 30, iter = Nsim, n_core=16, 
                                   file_cache = here::here("simulation", "p_inter_simple.rds"))
```

### FWER and CER

Type I error rates for main effects and interaction (when different correction methods were applied to the simple effect analysis):
```{r}
p_inter_simple %>% 
  select(iter, adjust, p_mainA, p_mainB, p_inter) %>% 
  distinct() %>% 
  group_by(adjust) %>% 
  summarize(mean(p_mainA<0.05), mean(p_mainB<0.05), mean(p_inter<0.05)) 
```

Type I error rates (for different combinations of simple effects and interaction):
```{r}
df_TypeI_inter <- sig_inter_simple(p_inter_simple, 0.05) %>% 
  select(-c(contrast, sig_simple)) %>% 
  distinct() %>% 
  group_by(adjust, group) %>% 
  summarize(TypeI_inter = mean(sig_inter),
            TypeI_any_simple = mean(sig_any_simple),
            `TypeI_inter&any` = mean(`sig_inter&any`),
            .groups="drop") %>% 
  pivot_longer(c(TypeI_any_simple, `TypeI_inter&any`), 
               names_to = "effects", values_to = "TypeI") %>% 
  mutate(effects = paste(effects, group, sep="_")) 

# Type I error rates
df_TypeI_inter %>% 
  select(-group) %>% 
  pivot_wider(names_from = effects, values_from = TypeI) %>% 
  select(adjust, TypeI_inter, TypeI_any_simple_byAB_4, 
         `TypeI_inter&any_byAB_4`, everything()) %>% 
  filter(adjust != "uncorrected") 
```

```{r}
df_TypeI_inter %>% 
  select(-group) %>% 
  filter(adjust == "uncorrected") 
```

```{r fig.width=10.5, fig.asp=0.35}
df_inter_simp_tmp <- df_TypeI_inter %>% 
  filter(group != "byB_2") %>% 
  select(-c(group, TypeI_inter)) %>% 
  add_row(adjust="uncorrected", effects = "Interaction\n", 
          TypeI = unique(df_TypeI_inter$TypeI_inter)) %>% 
  mutate(
    adjust = if_else(adjust == "uncorrected", "uncorrected",
                     str_to_title(as.character(adjust))),
    adjust = fct_reorder(adjust, TypeI, mean, .desc=TRUE),
    adjust = fct_relevel(adjust, "uncorrected"),
    effects = factor(effects, 
                     levels = c("Interaction\n", 
                                "TypeI_any_simple_byA_2", "TypeI_any_simple_byAB_4",
                                "TypeI_inter&any_byA_2", "TypeI_inter&any_byAB_4")),
    effects = fct_recode(
      effects, 
      `Simple effect analysis\n(two tests, FWER)`= "TypeI_any_simple_byA_2", 
      `Simple effect analysis\n(four tests, FWER)`="TypeI_any_simple_byAB_4",
      `Interaction & simple effects\nwith two tests (CER)`="TypeI_inter&any_byA_2",
      `Interaction & simple effects\nwith four tests (CER)`= "TypeI_inter&any_byAB_4"))

fig_simu_c <- df_inter_simp_tmp %>% 
  ggplot(aes(x=as.numeric(adjust), y = TypeI*100, fill=adjust)) + 
  geom_col(width=1) +   
  facet_grid(cols=vars(effects), scales="free_x", space="free_x", switch="x") +
  ggh4x::facetted_pos_scales(
    list(
      scale_x_continuous(breaks = 1, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65))
    )) + 
  scale_fill_manual(values = custom_clr[c(1,3:6)],
                    breaks = levels(df_inter_simp_tmp$adjust)) + #,
  geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
  scale_y_continuous(limits = c(0, 18), expand = c(0, 0)) +
  labs(y = "Type I Error Rate (%)", fill = "") +
  theme(legend.position = c(0.84, 0.65),
        axis.ticks.length = unit(.15, "cm"),
        axis.text.x = element_blank(),
        panel.spacing.x = unit(0, "pt"),
        strip.placement = "outside",
        axis.title.x = element_blank())
  
# ggsave(file.path("figures","fig_simu_TypeI_inter.png"), fig_simu_c, width = 10, height = 4)
fig_simu_c
```


```{r fig.width=10.5, fig.asp=0.35}
df_inter_simp_tmp_supp <- df_TypeI_inter %>% 
  filter(group != "byA_2") %>% # the only difference from previous figure
  select(-c(group, TypeI_inter)) %>% 
  add_row(adjust="uncorrected", effects = "Interaction\n", 
          TypeI = unique(df_TypeI_inter$TypeI_inter)) %>% 
  mutate(
    adjust = if_else(adjust == "uncorrected", "uncorrected",
                     str_to_title(as.character(adjust))),
    adjust = fct_reorder(adjust, TypeI, mean, .desc=TRUE),
    adjust = fct_relevel(adjust, "uncorrected"),
    effects = factor(effects, 
                     levels = c("Interaction\n", 
                                "TypeI_any_simple_byB_2", "TypeI_any_simple_byAB_4",
                                "TypeI_inter&any_byB_2", "TypeI_inter&any_byAB_4")),
    effects = fct_recode(
      effects, 
      `Simple effect analysis\n(two tests, FWER)`= "TypeI_any_simple_byB_2", 
      `Simple effect analysis\n(four tests, FWER)`="TypeI_any_simple_byAB_4",
      `Interaction & simple effects\nwith two tests (CER)`="TypeI_inter&any_byB_2",
      `Interaction & simple effects\nwith four tests (CER)`= "TypeI_inter&any_byAB_4"))

fig_simu_c_supp <- df_inter_simp_tmp_supp %>% 
  ggplot(aes(x=as.numeric(adjust), y = TypeI*100, fill=adjust)) + 
  geom_col(width=1) +   
  facet_grid(cols=vars(effects), scales="free_x", space="free_x", switch="x") +
  ggh4x::facetted_pos_scales(
    list(
      scale_x_continuous(breaks = 1, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65)),
      scale_x_continuous(breaks = 3, expand = c(0, 0.65))
    )) + 
  scale_fill_manual(values = custom_clr[c(1,3:6)],
                    breaks = levels(df_inter_simp_tmp$adjust)) + #,
  geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
  scale_y_continuous(limits = c(0, 18), expand = c(0, 0)) +
  labs(y = "Type I Error Rate (%)", fill = "") +
  theme(legend.position = c(0.84, 0.65),
        axis.ticks.length = unit(.15, "cm"),
        axis.text.x = element_blank(),
        panel.spacing.x = unit(0, "pt"),
        strip.placement = "outside",
        axis.title.x = element_blank())
  
# ggsave(file.path("figures","fig_simu_TypeI_inter_supp.png"), fig_simu_c_supp, width = 10, height = 4)
fig_simu_c_supp
```

### Single tests

```{r}
# with significant interaction and uncorrected simple effects
df_sig_inter_single1 <- sig_inter_simple(p_inter_simple, 0.05) %>% 
  filter(adjust == "uncorrected") %>% 
  transmute(Method = "Inter. & simple\n(uncorrected)",
            group, contrast, 
            sig_cber = `sig_inter&any`,
            sig_cber_single = sig_cber & sig_simple,
            sig_cber_inter = sig_cber & sig_inter) 

# with corrected simple effects
df_TypeI_inter_single <- sig_inter_simple(p_inter_simple, 0.05) %>% 
  filter(adjust != "uncorrected") %>% 
  transmute(Method = paste0(str_to_title(adjust), "\n(post-hoc only)"),
            group, contrast, 
            sig_cber = sig_any_simple,
            sig_cber_single = sig_cber & sig_simple,
            sig_cber_inter = sig_cber & sig_inter) %>% 
  # combine the two df and make it to long format
  bind_rows(df_sig_inter_single1) %>% 
  group_by(group, contrast, Method) %>% 
  summarize(TypeI_cber = mean(sig_cber),
            TypeI_cber_single = mean(sig_cber_single),
            TypeI_cber_inter = mean(sig_cber_inter),
            .groups="drop") %>% 
  pivot_wider(c(group, Method, TypeI_cber, TypeI_cber_inter), 
              names_from = contrast, values_from = TypeI_cber_single) %>% 
  pivot_longer(!c("group", "Method"), 
               names_to = "contrast", values_to = "TypeI") %>% 
  filter(!is.na(TypeI)) %>% 
  # update the level names and order
  mutate(contrast = case_when(
    contrast == "a1 b1 - a1 b2" ~ "a1_(b1 - b2)",
    contrast == "a1 b1 - a2 b1" ~ "b1_(a1 - a2)",
    contrast == "a1 b2 - a2 b2" ~ "b2_(a1 - a2)",
    contrast == "a2 b1 - a2 b2" ~ "a2_(b1 - b2)",
    contrast == "TypeI_cber" ~ "CBER",
    contrast == "TypeI_cber_inter" ~ "Interaction",
    TRUE ~ contrast),
    contrast = as_factor(contrast),
    contrast = fct_relevel(contrast, "CBER", "Interaction",
                           "a1_(b1 - b2)", "a2_(b1 - b2)"),
    Method = fct_reorder(Method, TypeI, mean, .desc=TRUE),
    Method = fct_relevel(Method, "Inter. & simple\n(uncorrected)"),
    grouplabel = if_else(group=="byAB_4", 
                         "Simple effect analysis with four tests",
                         "Simple effect analysis with two tests"),
    grouplabel = as_factor(grouplabel),
    grouplabel = fct_relevel(grouplabel, "Simple effect analysis with two tests")
  ) %>% 
  filter(contrast!="Interaction")

df_TypeI_inter_single
```

```{r fig.width=11.5, fig.asp=0.5}
fig_simu_single_c <- df_TypeI_inter_single %>% 
  filter(group != "byB_2") %>% 
  ggplot(aes(x = Method, y = TypeI*100, fill = contrast, group = contrast)) +
  geom_col(position = "dodge") +
  facet_grid(cols=vars(grouplabel), switch="x") +
  scale_fill_manual(values = custom_clr) +
  geom_hline(yintercept = 5, linetype= "dashed", color="gray25") +
  scale_y_continuous(limits = c(0, 7), expand = c(0, 0), breaks = 1:6) +
  labs(y = "Type I Error Rate (%)", fill = "") +
  guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
  theme(legend.position = c(0.5, 0.87),
        legend.direction="horizontal",
        axis.ticks.length = unit(.15, "cm"),
        strip.placement = "outside",
        axis.title.x = element_blank())

# ggsave(file.path("figures","fig_simu_single_inter.png"), fig_simu_single_c, width = 12, height = 5)
fig_simu_single_c
```


# Publication 

## Figure 2

```{r fig.width=10.5, fig.asp=0.7}
fig_simu <- ggarrange(fig_simu_a, fig_simu_b,
          nrow = 1, labels = c("a", "b"),
          widths = c(0.35, 0.65)) %>% 
  ggarrange(fig_simu_c,
            nrow = 2, labels = c("", "c"))

# ggsave(file.path("figures","fig_simu.png"), fig_simu, width = 12, height = 8.4)
fig_simu
```

```{r fig.width=12, fig.asp=0.85}
fig_simu_single <- ggarrange(fig_simu_single_a, fig_simu_single_b,
          nrow = 1, labels = c("a", "b"),
          widths = c(0.3, 0.7)) %>% 
   ggarrange(fig_simu_single_c,
            nrow = 2, labels = c("", "c"))

# ggsave(file.path("figures","fig_simu_single.png"), fig_simu_single, width = 12, height = 10)
fig_simu_single
```


## Table 1


# Supplemenatry

```{r fig.width=13, fig.asp=0.4}
fig_simu_a_supp
```


```{r fig.width=8, fig.asp=0.8}
fig_simu_single_a_supp
```


```{r fig.width=13, fig.asp=.7}
fig_simu_b_supp
```

```{r fig.width=14, fig.asp=1.1}
fig_simu_single_b_supp
```

```{r fig.width=10.5, fig.asp=0.35}
fig_simu_c_supp
```

# Session information {.unnumbered}
```{r}
sessionInfo()
```




