---
title             : "Figure 1 used in rethink ANOVA"

author: 
  - name: "Haiyang Jin"
    affiliation: ""
    address: ""
    email: "haiyang.jin@outlook.com"
    corresponding: yes
    

authornote        : |
  This is the code for creating the Figure 1 in the rethinking ANOVA paper.
        
date: "`r format(Sys.time(), '%d-%m-%Y')`"

---

```{r setup and load the related libraries, appendix, include=FALSE}
## load libraries
library(knitr)
library(tidyverse)
library(faux)
library(afex)
library(emmeans)
library(ggpubr)
library(here)
library(papaja)

theme_set(theme_apa())

# set global chunk options, put figures into folder
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(
  echo = FALSE,
  include = TRUE,
  warning = FALSE,
  replace.assign=TRUE,
  width = 1800, 
  message = FALSE,
  fig.align='center'
)

set.seed(2022)

# source(here("R", "funcs.R"))

```

# Interaction and simple effect analysis

## Sig interaction with sig simple effect
```{r}
set.seed(2021)
df_simu_ss <-
  add_random(subj=1:60) %>% 
  add_between(group=c("control", "train"), .by="subj") %>% 
  add_within(test = c("pre", "post")) %>% 
  add_contrast("group", "treatment", colnames = "group_dummy") %>% 
  add_contrast("test", "treatment", colnames = "test_dummy") %>% 
  add_ranef("subj", s_int=.1, s_slp=.1, .cors=.5) %>% 
  add_ranef(sigma=3) %>% 
  mutate(d = (2 + s_int) + group_dummy*.5 + (test_dummy + s_slp)*.5 +
           group_dummy * (test_dummy + s_slp)*1.5 + sigma)
  
str(df_simu_ss)
```

```{r}
aov_simu_ss <- aov_4(d ~ group * test + (test | subj),
                   data = df_simu_ss)
aov_simu_ss
```

```{r}
emm_simu_ss <- emmeans(aov_simu_ss, ~ group + test)
(contr_simu_ss <- contrast(emm_simu_ss, "pairwise", adjust="none"))[c(1,2,5,6)]
```

```{r}
emmip(emm_simu_ss, group ~ test, CIs = T)
```

## Sig interaction with non-sig simple effect
Significant interaction with non-significant simple effect analysis results:

```{r}
set.seed(2021)

df_simu_sn <-
  add_random(subj=1:60) %>% 
  add_between(group=c("control", "train"), .by="subj") %>% 
  add_within(test = c("pre", "post")) %>% 
  add_contrast("group", "treatment", colnames = "group_dummy") %>% 
  add_contrast("test", "treatment", colnames = "test_dummy") %>% 
  add_ranef("subj", s_int=.1, s_slp=.1, .cors=.5) %>% 
  add_ranef(sigma=3) %>% 
  mutate(d = (1.8 + s_int) + group_dummy*3 + (test_dummy + s_slp)*1.5 +
           group_dummy * (test_dummy + s_slp)*(-3.3) + sigma)

str(df_simu_sn)
```

```{r}
aov_simu_sn <- aov_4(d ~ group * test + (test | subj),
                   data = df_simu_sn)
aov_simu_sn
```

```{r}
emm_simu_sn <- emmeans(aov_simu_sn, ~ group + test)
(contr_simu_sn <- contrast(emm_simu_sn, "pairwise", adjust="none"))[c(1,2,5,6)]
```

```{r}
emmip(emm_simu_sn, group ~ test, CIs = T)
```

## Non-sig interaction with sig simple effect
Non-significant interaction with significant simple effect analysis results:
```{r}
set.seed(2021)

df_simu_ns <-
  add_random(subj=1:60) %>% 
  add_between(group=c("control", "train"), .by="subj") %>% 
  add_within(test = c("pre", "post")) %>% 
  add_contrast("group", "treatment", colnames = "group_dummy") %>% 
  add_contrast("test", "treatment", colnames = "test_dummy") %>% 
  add_ranef("subj", s_int=.1, s_slp=.1, .cors=.5) %>% 
  add_ranef(sigma=3) %>% 
  mutate(d = (1.5 + s_int) + group_dummy*1 + (test_dummy + s_slp)*1.5 +
           group_dummy * (test_dummy + s_slp)*.3 + sigma)

str(df_simu_ns)
```

```{r}
aov_simu_ns <- aov_4(d ~ group * test + (test | subj),
                   data = df_simu_ns)
aov_simu_ns
```

```{r}
emm_simu_ns<- emmeans(aov_simu_ns, ~ group + test)
(contr_simu_ns <- contrast(emm_simu_ns, "pairwise", adjust="none"))[c(1,2,5,6)]
```

```{r}
emmip(emm_simu_ns, group ~ test, CIs = T)
```

## Non-sig interaction with non-sig simple effect
Non-significant interaction with non-significant simple effect analysis results:
```{r}
set.seed(2021)

df_simu_nn <- 
  add_random(subj=1:60) %>% 
  add_between(group=c("control", "train"), .by="subj") %>% 
  add_within(test = c("pre", "post")) %>% 
  add_contrast("group", "treatment", colnames = "group_dummy") %>% 
  add_contrast("test", "treatment", colnames = "test_dummy") %>% 
  add_ranef("subj", s_int=.1, s_slp=.1, .cors=.5) %>% 
  add_ranef(sigma=3) %>% 
  mutate(d = (2.5 + s_int) + group_dummy*1.3 + (test_dummy + s_slp)*(-.5) +
           group_dummy * (test_dummy + s_slp)*(-.5) + sigma)

str(df_simu_nn)
```

```{r}
aov_simu_nn <- aov_4(d ~ group * test + (test | subj),
                   data = df_simu_nn)
aov_simu_nn
```

```{r}
emm_simu_nn<- emmeans(aov_simu_nn, ~ group + test)
(contr_simu_nn <- contrast(emm_simu_nn, "pairwise", adjust="none"))[c(1,2,5,6)]
```

```{r}
emmip(emm_simu_nn, group ~ test, CIs = T)
```

## Combined plots

```{r}

df_inter <- rbind(
  mutate(as_tibble(aov_simu_ss$anova_table)[3,],
         interaction = "significant", `simple effect` = "significant",
         x = 1.5, y = 1.5),
  mutate(as_tibble(aov_simu_sn$anova_table)[3,],
         interaction = "significant", `simple effect` = "nonsignificant",
         x = 1.5, y = 1.5),
  mutate(as_tibble(aov_simu_ns$anova_table)[3,],
         interaction = "nonsignificant", `simple effect` = "significant",
         x = 1.5, y = 1.5),
  mutate(as_tibble(aov_simu_nn$anova_table)[3,],
         interaction = "nonsignificant", `simple effect` = "nonsignificant",
         x = 1.5, y = 1.5)
) %>% 
  mutate(group = "control1", test = "pre1",
         ptext = if_else(`Pr(>F)` < .001, paste0("italic(p)[inter] < .001"), 
                         paste0("italic(p)[inter] == ", round(`Pr(>F)`,3))),
         interaction = fct_relevel(interaction, "significant", "nonsignificant"),
         `simple effect` = fct_relevel(`simple effect`, "significant", "nonsignificant"))

df_simp <- rbind(
  mutate(filter(as_tibble(contr_simu_ss)[c(1,2,5,6)], p.value==min(p.value)),
         interaction = "significant", `simple effect` = "significant",
         x = 1.5, y = 1),
  mutate(filter(as_tibble(contr_simu_sn)[c(1,2,5,6)], p.value==min(p.value)),
         interaction = "significant", `simple effect` = "nonsignificant",
         x = 1.5, y = 1),
  mutate(filter(as_tibble(contr_simu_ns)[c(1,2,5,6)], p.value==min(p.value)),
         interaction = "nonsignificant", `simple effect` = "significant",
         x = 1.5, y = 1),
  mutate(filter(as_tibble(contr_simu_nn)[c(1,2,5,6)], p.value==min(p.value)),
         interaction = "nonsignificant", `simple effect` = "nonsignificant",
         x = 1.5, y = 1)
) %>% 
  mutate(group = "control1", test = "pre1",
         labeltext_inter = paste0(interaction, "~interaction"),
         labeltext_sim = paste0(`simple effect`, "~simple~effects"),
         ptext = if_else(`p.value` < .001, paste0("italic(p)[min_sim] < .001"), 
                         paste0("italic(p)[min_sim] == ", round(`p.value`,3))),
         interaction = fct_relevel(interaction, "significant", "nonsignificant"),
         `simple effect` = fct_relevel(`simple effect`, "significant", "nonsignificant"))

```

```{r fig.asp=.7, fig.width=7}
theme_set(theme_apa())

rbind(mutate(as_tibble(emm_simu_ss),
             interaction = "significant", `simple effect` = "significant"),
      mutate(as_tibble(emm_simu_sn),
             interaction = "significant", `simple effect` = "nonsignificant"),
      mutate(as_tibble(emm_simu_ns),
             interaction = "nonsignificant", `simple effect` = "significant"),
      mutate(as_tibble(emm_simu_nn),
             interaction = "nonsignificant", `simple effect` = "nonsignificant")
) %>% 
  mutate(test = fct_recode(test, "pre-test" = "pre", "post-test" = "post"),
         interaction = fct_relevel(interaction, "significant", "nonsignificant"),
         `simple effect` = fct_relevel(`simple effect`, "significant", "nonsignificant")) %>% 
  ggplot(aes(test, emmean, group=group, color=group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15, 
                position=position_dodge(.1), show.legend=F) +
  facet_grid(`simple effect` ~ interaction, labeller = label_both) +
  geom_text(aes(label=ptext, x=x, y=y), data=df_inter,
            color="black", size = 3, show.legend = F, parse=T) +
  geom_text(aes(label=ptext, x=x, y=y), data=df_simp,
            color="black", size = 3, show.legend = F, parse=T) +
  scale_color_manual(values=c("#56B4E9", "#D55E00")) +
  labs(x=NULL, y=expression("Sensitivity"~italic("d'")), 
       color="Group", linetype="Group") +
  theme(legend.position=c(.5, .5)) +
  NULL

ggsave(file="inter_simple_sig_train.png", width = 7.2, height = 5.4)
```

```{r fig.asp=.7, fig.width=3}
theme_set(theme_apa())

rbind(mutate(as_tibble(emm_simu_ss),
             interaction = "significant", `simple effect` = "significant"),
      mutate(as_tibble(emm_simu_sn),
             interaction = "significant", `simple effect` = "nonsignificant"),
      mutate(as_tibble(emm_simu_ns),
             interaction = "nonsignificant", `simple effect` = "significant"),
      mutate(as_tibble(emm_simu_nn),
             interaction = "nonsignificant", `simple effect` = "nonsignificant")
) %>% 
  mutate(interaction = fct_relevel(interaction, "significant", "nonsignificant"),
         `simple effect` = fct_relevel(`simple effect`, "significant", "nonsignificant")) %>% 
  ggplot(aes(test, emmean, group=group, color=group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15, 
                position=position_dodge(.1), show.legend=F) +
  facet_grid(`simple effect` ~ interaction) +
  geom_text(aes(label=labeltext_inter), data=df_simp, x=.5, y = 4.5,
            color="black", size = 3, hjust=0, show.legend = F, parse=T) +
  geom_text(aes(label=labeltext_sim), data=df_simp, x=.5, y = 4.3,
            color="black", size = 3, hjust=0, show.legend = F, parse=T) +
  geom_text(aes(label=ptext, x=x, y=y), data=df_inter,
            color="black", size = 3, show.legend = F, parse=T) +
  geom_text(aes(label=ptext, x=x, y=y), data=df_simp,
            color="black", size = 3, show.legend = F, parse=T) +
  labs(x="Alignment", y=expression("Sensitivity"~italic("d'")),
       color="Congruency", linetype="Congruency") +
  ylim(c(2.25, 4.45)) +
  # theme_apa_nostrip() +
  # theme(legend.position="bottom") +
  theme(legend.position=c(.5,.55), legend.direction="horizontal") +
  NULL

# ggsave(file=here::here("images", "inter_simple_sig.png", width = 7.2, height = 5.4)
```
